---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Top 100 U.S. Climate, Air, and Water Polluters (data from 2018)

<!-- badges: start -->
<!-- badges: end -->

## Dataset Description
Data on the combined toxic 100 / greenhouse 100 indexes (2020 report, based on 2018 data). 

Researchers at the Political Economy Research Institute (PERI) at the University of Massachusetts Amherst published new editions of the Greenhouse 100 Index, Toxic 100 Air Index, and the Toxic 100 Water Index. The Greenhouse 100 Index ranks U.S. companies by their emissions from their large facilities responsible for global climate change. They used data from the U.S. EPA Greenhouse Gas Reporting Program to determine the greenhouse ranks. The Toxic 100 Air and Toxic 100 Water Indexes rank U.S. industrial polluters by the toxics from their large facilities. They used data from the U.S. EPA Toxics Release Inventory to determine the toxic ranks. 

The PERI Indexes also includes Environmental Justice indicators to assess impacts on people below the poverty line and racial/ethnic minorities. Specifically, these Environmental Justice variables measure marginalized groups possible exposure to the toxins released by specific corporations/entities, and the percent of the total at-risk population each subgroup makes up.

## Variable Description

### topPolluters
- A data frame with 220 observations on the following 8 variables:

### polluter
- Parent Entity or Corporation

### toxic.air.rank
- Toxic 100 Air Rank; the company's rank based on air toxics from large facilities in 2018

### greenhouse.rank
- Greenhouse 100 Rank; the company's rank based on greenhouse gas emissions from large facilities in 2018

### toxic.water.rank
- Toxic 100 Water Rank; the company's rank based on water toxics from large facilities in 2018

### toxic.air.poor
- Toxic 100 Air EJ: Poor Share; shares of potential exposure to air toxics borne by people living below the poverty line

### toxic.air.minority
- Toxic 100 Air EJ: Minority Share; shares of potential exposure to air toxics borne by people in minority racial/ethnic groups

### greenhouse.poor
- Greenhouse 100 EJ: Poor Share; shares of potential exposure to co-pollutants of combustion borne by people living below the poverty line

### greenhouse.minority
- Greenhouse 100 EJ: Minority Share; shares of potential exposure to co-pollutants of combustion borne by people in minority racial/ethnic groups

### Source
- https://www.peri.umass.edu/combined-toxic-100-greenhouse-100-indexes-current

## Installation

```{r, echo=FALSE}
install.packages("devtools")
install.packages("topPolluters")
library(topPolluters)
library(tidyr)
library(ggplot2)
library(viridis)
library(forcats)
library(knitr)
### Sample of the data.
kable(head(topPolluters))
```
## Use
Environmental racism and discrimination is very pertinent in today's world as it is a huge contributor to various health issues, and this dataset can help answer important questions regarding what U.S. corporations are contributing to different types of pollution both overall and in different communities.

### Example 1 - Are there any corporations/entities that rank highly (as top polluters) in all three indexes?
```{r}
filter(topPolluters, toxic.air.rank <= 50, greenhouse.rank <= 50, toxic.water.rank <= 50) %>%
  pivot_longer(
    cols = c(toxic.air.rank, greenhouse.rank, toxic.water.rank),
    names_to = "rank.type",
    values_to = "ranking") %>%
  ggplot(aes(x = polluter, y = ranking, label = ranking, colour = rank.type)) +
  geom_linerange(aes(x = polluter, ymin = 50, ymax = ranking, colour = rank.type), 
                   position = position_dodge(width = .4))+
  geom_point(aes(x = polluter, y = ranking, colour = rank.type),
               position = position_dodge(width = .4))+
  scale_x_discrete(expand = expansion(add = c(.3, .3))) +
  scale_y_continuous(trans = "reverse",
                     expand = expansion(add = c(.6, 4))) +
  scale_color_manual(
    labels = c("Greenhouse Rank", "Toxic Air Rank", "Toxic Water Rank"),
    values = c("#11B31A", "#D8BC42", "#4DA2F0")) +
  geom_text(aes(label = round(ranking, .4)), 
            position = position_dodge(0.4), vjust = -.4)+
  labs( 
    title = "U.S. Corporations/Entities that are Top Climate, Air, and Water Polluters",
    subtitle = "All Corporations/Entities are in the top 50 of the Greenhouse 100 Index,\nToxic 100 Air Polluters Index, and Toxic 100 Water Polluters Index",
    y = "Ranking on Each Index\n(the smaller the number, the worse the polluter)"
  ) +
  theme_light()+
  theme(legend.position = "top",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = -90, hjust = 0, vjust = .5)) 
```

### Example 2 - Do top toxic air polluters negatively impact marginalized communities at a disproportionate rate?
```{r}
topPolluters %>%
  filter(toxic.air.rank <= 30) %>%
  select(polluter, toxic.air.rank, toxic.air.poor, toxic.air.minority) %>%
  pivot_longer(
    cols = c(toxic.air.minority, toxic.air.poor),
    names_to = "metric",
    values_to = "percentage"
  ) %>%
  mutate(percentage = percentage * 100) %>%
  ggplot(aes(x = polluter, y = percentage, fill = metric)) +
  geom_bar(width = 0.7, position = "dodge", stat = "identity") +
  theme_bw() +
  aes(x = fct_reorder(polluter, toxic.air.rank, .desc = TRUE)) +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80, 100),
    labels = c("0%", "20%", "40%", "60%", "80%", "100%"),
    expand = expansion(add = c(0, 2))
  ) +
  labs(
    title = "The Top 30 U.S. Industrial Pollutors of Air Toxics of 2018\nand the Resulting Exposure to Low-Income People and\nRacial/Ethnic Minorities",
    caption = "The ranking of the top air pollutors is based on total potential chronic human health risk from their facilities.\nThe percentages are in regards to the makeup of the at-risk populations (which are typically those nearest\nto the toxic facilities), but spefically how much of the at risk population is below the poverty line or is a\nracial/ethnic minority.",
    x = "Top Pollutors of Air Toxics\n(in Descending Order of Rank)",
    y = "Percent of Total At-Risk Population"
  ) +
  theme(
    legend.title = element_blank(),
    plot.title = element_text(hjust = 0),
    plot.caption = element_text(hjust = 0, size = 8),
    legend.position = "top"
  ) +
  scale_fill_manual(
    values = c("#E69F00", "#56B4E9"),
    breaks = c("toxic.air.poor", "toxic.air.minority"),
    labels = c("People below the poverty line", "Racial/ethnic minorities")
  ) +
  coord_flip()
```

