---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Top Climate, Air, and Water Polluters (2018)

<!-- badges: start -->
<!-- badges: end -->

Researchers at the Political Economy Research Institute (PERI) at the University of Massachusetts Amherst created this all encompassing dataset on the top 100 parent entities or corporations that contributed to greenhouse gas pollution, air pollution, and water pollution. 

## Installation

The development version of XXX is available from [GitHub](https://github.com/) with:
```{r}
install.packages("devtools")
install.packages("topPolluters")
```
## Example 1 - U.S. Companies that are in the Top 50 Climate, Air, and Water Polluter Indexs
```{r}
library(topPolluters)
library(tidyr)
library(ggplot2)
library(viridis)
library(forcats)
head(topPolluters)

filter(topPolluters, toxic.air.rank <= 50, greenhouse.rank <= 50, toxic.water.rank <= 50) %>%
  pivot_longer(
    cols = c(toxic.air.rank, greenhouse.rank, toxic.water.rank),
    names_to = "rank.type",
    values_to = "ranking") %>%
  ggplot(aes(x = polluter, y = ranking, label = ranking, colour = rank.type)) +
  geom_linerange(aes(x = polluter, ymin = 50, ymax = ranking, colour = rank.type), 
                   position = position_dodge(width = .4))+
  geom_point(aes(x = polluter, y = ranking, colour = rank.type),
               position = position_dodge(width = .4))+
  scale_x_discrete(expand = expansion(add = c(.3, .3))) +
  scale_y_continuous(trans = "reverse",
                     expand = expansion(add = c(.6, 4))) +
  scale_color_manual(
    labels = c("Greenhouse Rank", "Toxic Air Rank", "Toxic Water Rank"),
    values = c("#11B31A", "#D8BC42", "#4DA2F0")) +
  geom_text(aes(label = round(ranking, .4)), 
            position = position_dodge(0.4), vjust = -.4)+
  labs( 
    title = "U.S. Companies that are in the Top 50 Climate, Air, and Water Polluter Indexes",
    subtitle = "From PERI's Greenhouse 100 Index, Toxic 100 Air Polluters Index,\nand Toxic 100 Water Polluters Index",
    y = "Rank on Each Index"
  ) +
  theme_light()+
  theme(legend.position = "top",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = -90, hjust = 0)) 
```
## Example 2 - The Top 30 U.S. Industrial Pollutors of Air Toxics of 2018 and Resulting Exposure to Low-Income People and Racial/Ethnic Minorities
```{r}
topPolluters %>%
  filter(toxic.air.rank <= 30) %>%
  select(polluter, toxic.air.rank, toxic.air.poor, toxic.air.minority) %>%
  pivot_longer(
    cols = c(toxic.air.minority, toxic.air.poor),
    names_to = "metric",
    values_to = "percentage"
  ) %>%
  mutate(percentage = percentage * 100) %>%
  ggplot(aes(x = polluter, y = percentage, fill = metric)) +
  geom_bar(width = 0.7, position = "dodge", stat = "identity") +
  theme_bw() +
  aes(x = fct_reorder(polluter, toxic.air.rank, .desc = TRUE)) +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80, 100),
    labels = c("0%", "20%", "40%", "60%", "80%", "100%"),
    expand = expansion(add = c(0, 2))
  ) +
  labs(
    title = "The Top 30 U.S. Industrial Pollutors of Air Toxics of 2018 and the\nResulting Exposure to Low-Income People and Racial/Ethnic Minorities",
    caption = "The ranking of the top air pollutors is based on total potential chronic human health risk from their facilities.\nThe percentages are in regards to the makeup of the at-risk populations (which are typically those nearest to the toxic facilities),\nbut spefically how much of the at risk population is below the poverty line or is a racial/ethnic minority.",
    x = "Top Pollutors of Air Toxics (in Descending Order of Rank)",
    y = "Percent Group Makes Up of Total At-Risk Population"
  ) +
  theme(
    legend.title = element_blank(),
    plot.title = element_text(hjust = 0),
    plot.caption = element_text(hjust = 0, size = 8),
    legend.position = "top"
  ) +
  scale_fill_manual(
    values = c("#E69F00", "#56B4E9"),
    breaks = c("toxic.air.poor", "toxic.air.minority"),
    labels = c("People below the poverty line", "Racial/ethnic minorities")
  ) +
  coord_flip()
```

